Inference for Trans-dimensional Bayesian Models with Diffusive Nested Sampling

### 2015-01-14 Documents/Paper/ms.tex
png figures
Italic captions

\begin{figure}
\begin{center}
\includegraphics{pgm.pdf}
\caption{A probabilistic graphical model (PGM) depicting the kind
of model discussed in this paper. Produced using daft ({\tt daft-pgm.org}).
The number, $N$, of components in the model is an unknown parameter, as are the
properties of the components. The data depends on the properties of the components.
Note that it is possible (and common) to have other parameters in the model
that point directly to the data. However, we have omitted such parameters
for simplicity.
\label{fig:pgm}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{minipage}[lr]{\textwidth}
\begin{center}
\includegraphics[width=2.7in]{challenges.pdf}
\includegraphics[width=2.7in]{dnest.png}
\end{center}
\end{minipage}
\caption{{\bf Left}: An example of a two-dimensional posterior distribution
with multiple modes, strong dependence, and a phase transition (in the top
right mode). While this low dimensional example is not difficult to sample,
an analogous problem in high dimensions would be. {\bf Right}: Diffusive
Nested Sampling replaces the target posterior distribution by a mixture of
constrained priors. The density is non-negligible everywhere (since the prior
is a component of this distribution), and states intermediate between the two
phases in the top-right mode are appropriately up-weighted.
\label{fig:challenges}}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{sinewave_data.pdf}
\caption{The simulated data for the sinusoidal example. The solid line shows
the true signal which consists of a large slow oscillation and a much smaller
amplitude fast oscillation. The points are the measurements, simulated with a
noise standard deviation of $\sigma = 0.5$.
\label{fig:sinewave_data}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{trace_logl.pdf}
\caption{A ``trace plot'' of the log likelihood in the sinusoidal model, for an
MCMC chain designed to sample the posterior distribution. The two phases
have log-likelihoods around -730 and -745 respectively, however transitions
between the two phases are quite uncommon.
\label{fig:trace_logl}}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{sinewaves_likelihood.pdf}
\caption{{\bf Top panel: }
The shape of the likelihood function with respect to prior mass.
{\bf Bottom panel: }The posterior weights of the saved samples. The existence
of a phase transition causes this to display two separate peaks, which
would be difficult to mix between if we simply attempted to sample the posterior
distribution.
\label{fig:sinewaves_likelihood}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{N_result.pdf}
\caption{The inference for $N$, the number of sinusoids, based on the
sinusoid data. The true value was $N=2$, which is also the most probable value
given the data. Of course, this is not always the case.
\label{fig:N_result}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{trace_logl.pdf}
\caption{A ``trace plot'' of the log likelihood in the SineWave model, for an
MCMC chain designed to sample the posterior distribution. The two phases
have log-likelihoods around -730 and -745 respectively, however transitions
between the two phases are quite uncommon.
\label{fig:trace_logl}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.7]{galaxyfield_data.pdf}
\caption{The simulated image for the GalaxyField example. The image is
$200\times 200$ pixels in size and contains $N=47$ ``galaxies''.
\label{fig:galaxyfield_data}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{galaxyfield_likelihood.pdf}
\caption{{\bf Top panel: }
The shape of the likelihood function with respect to prior mass for the
GalaxyField data. Note the phase
transition at $\log(X) \approx -100$ nats. This separates models with just
the few brightest galaxies from models with many faint galaxies as well.
While this doesn't affect the
posterior distribution (unlike the sinewave example), it would affect calculation
of the marginal likelihood if annealing were used.
{\bf Bottom panel: }The posterior weights of the saved samples.
\label{fig:galaxyfield_likelihood}}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{N_galaxy_result.pdf}
\caption{The inference for $N$, the number of galaxies, based on the
GalaxyField data. The true value was $N=47$.
\label{fig:N_galaxy_result}}
\end{center}
\end{figure}

>>>>

\begin{figure}
\begin{center}
\includegraphics{pgm.pdf}
\caption{\it A probabilistic graphical model (PGM) depicting the kind
of model discussed in this paper. Produced using daft ({\tt daft-pgm.org}).
The number, $N$, of components in the model is an unknown parameter, as are the
properties of the components. The data depends on the properties of the components.
Note that it is possible (and common) to have other parameters in the model
that point directly to the data. However, we have omitted such parameters
for simplicity.
\label{fig:pgm}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{minipage}[lr]{\textwidth}
\begin{center}
\includegraphics[width=2.7in]{challenges.pdf}
\includegraphics[width=2.7in]{dnest.png}
\end{center}
\end{minipage}
\caption{\it {\bf Left}: An example of a two-dimensional posterior distribution
with multiple modes, strong dependence, and a phase transition (in the top
right mode). While this low dimensional example is not difficult to sample,
an analogous problem in high dimensions would be. {\bf Right}: Diffusive
Nested Sampling replaces the target posterior distribution by a mixture of
constrained priors. The density is non-negligible everywhere (since the prior
is a component of this distribution), and states intermediate between the two
phases in the top-right mode are appropriately up-weighted.
\label{fig:challenges}}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{sinewave_data.pdf}
\caption{\it The simulated data for the sinusoidal example. The solid line shows
the true signal which consists of a large slow oscillation and a much smaller
amplitude fast oscillation. The points are the measurements, simulated with a
noise standard deviation of $\sigma = 0.5$.
\label{fig:sinewave_data}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{trace_logl.pdf}
\caption{\it A ``trace plot'' of the log likelihood in the sinusoidal model, for an
MCMC chain designed to sample the posterior distribution. The two phases
have log-likelihoods around -730 and -745 respectively, however transitions
between the two phases are quite uncommon.
\label{fig:trace_logl}}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.6]{sinewaves_likelihood.png}
\caption{\it {\bf Top panel: }
The shape of the likelihood function with respect to prior mass.
{\bf Bottom panel: }The posterior weights of the saved samples. The existence
of a phase transition causes this to display two separate peaks, which
would be difficult to mix between if we simply attempted to sample the posterior
distribution.
\label{fig:sinewaves_likelihood}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{N_result.pdf}
\caption{\it The inference for $N$, the number of sinusoids, based on the
sinusoid data. The true value was $N=2$, which is also the most probable value
given the data. Of course, this is not always the case.
\label{fig:N_result}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{trace_logl.pdf}
\caption{\it A ``trace plot'' of the log likelihood in the SineWave model, for an
MCMC chain designed to sample the posterior distribution. The two phases
have log-likelihoods around -730 and -745 respectively, however transitions
between the two phases are quite uncommon.
\label{fig:trace_logl}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.7]{galaxyfield_data.pdf}
\caption{\it The simulated image for the GalaxyField example. The image is
$200\times 200$ pixels in size and contains $N=47$ ``galaxies''.
\label{fig:galaxyfield_data}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.6]{galaxyfield_likelihood.png}
\caption{\it {\bf Top panel: }
The shape of the likelihood function with respect to prior mass for the
GalaxyField data. Note the phase
transition at $\log(X) \approx -100$ nats. This separates models with just
the few brightest galaxies from models with many faint galaxies as well.
While this doesn't affect the
posterior distribution (unlike the sinewave example), it would affect calculation
of the marginal likelihood if annealing were used.
{\bf Bottom panel: }The posterior weights of the saved samples.
\label{fig:galaxyfield_likelihood}}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{N_galaxy_result.pdf}
\caption{\it The inference for $N$, the number of galaxies, based on the
GalaxyField data. The true value was $N=47$.
\label{fig:N_galaxy_result}}
\end{center}
\end{figure}
